### 2012-10-09 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
tidying up

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}
%\date{}

@@

@@

\subsection{My view}
On the second consideration, 
we take for granted that there is a parallel between joint action and ordinary, individual action.
But is the standard view about the relation between action and intention correct?

Not everyone accepts the standard view.
Some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' \citet[p.\ 251]{bratman:2000_valuing}.%
%\footnote{
%For instance, \citet[p.\ 251]{bratman:2000_valuing} % = \citep[p.\ 51]{Bratman:2007bd}
%describes an agent called `Creature 2' who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.
%}

@@

\section{Intention and Motor Representation}
[\emph{*Probably skip all of this section in the talk}]

@@

Motor representations are relevantly similar to intentions,
as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
Of course motor representations differ from intentions in some important ways (as these authors note).
But they are similar in the respects that matter for explaining the purposiveness of action.
For, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
So anyone who accepts the standard story about purposive action and intention
should also accept that a similar story about purposive action and motor representation.
Given that the two basic questions about ordinary, individual purposive action can also be answered by appeal to intention,
they can also be answering by appeal to motor representation.
This motivates the following view.
Not all purposive actions are bound together and linked to outcomes by intentions. 
In some cases what binds together purposive actions and links them to outcomes are motor representations.
%In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
And of course in many cases it may be that both intention and motor representation are involved.

This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining what action is (just as intention does).
Now let’s return to joint action.

@@

This needs unpacking. 
A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.
One consequence is that motor representations are not concerned with merely kinematic or dynamic features of actions only.
Rather, some motor representations represent outcomes, such as the movement of a target object from one place to another.

@@

Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome. 
It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
What is much more controversial is that reciprocal motor representation occurs in joint action.
Some of the research that Gunether Knoblich presented supports the claim that there are reciprocal motor representations in joint action, and he suggests that these enable joint action.%
\footnote{
See \citet{kourtis:2012_predictive}: `the partner’s expected action is simulated at the motor level, which probably facilitates effective performance of the joint action.'
\citet{kourtis:2010_favoritism} show that 
reciprocal motor representation is more likely to occur in joint action than is mere observation. 
See also \citet{Knoblich:2003nf}.
 }

I want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of all of the agents.
So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions.
For example, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
In this case I may represent the movement of the object from A to B and not only the component movements.
I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
%
%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.

@@

What follows is speculative philosophy: we take for granted that sometimes social motor representation facilitates joint action and ask whether this conjecture bears on our question about which events are joint actions.

@@

To answer this question
let us take a step back and consider an individual action.
Suppose an agent moves a mug from one place to another, passing it from her left hand to her right hand half way.
It is a familiar idea that motor planning, like planning generally, involves starting with relatively abstract representations of outcomes and gradually filling in details.
We can capture this by supposing that 
motor representations for planning and monitoring action involve a hierarchical structure of representations.
At the top we might find a relatively abstract representation of an outcome, in this case of the movement of the object from one location to another.
Action-relevant details are progressively filled in by representations at lower stages of the hierarchy. 
Now in the action we are considering there is a need, even for the single agent, to coordinate the exchange between the two hands.
How is this achieved? 
We suppose that part of the answer involves the fact that planning for the movements of each hand is not done entirely independently.
Rather there is a plan for the whole action
and plans for the movements of each hand are components of this larger plan.
It is in part because they are parts of a larger plan that the plan for one hand constrains and is constrained by the plan for the other hand.

How is this relevant to the case of joint action?
In joint action the agents have the same goal, to move the object from one place to another.
They also face a similar coordination problem, requiring a precisely timed swap from one hand to another.
Now suppose, 
inspired by Koutis et al’s and others' findings,
that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
The planning is the same almost up to the actual muscle contractions.

How could this be helpful?
Suppose the agents' planning processes are similar enough that, for a given context and problem, they will produce approximately the same plans.
Then having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action,
and 
(iii) each  agent's plans for their own action are constrained by their plans for the other agent's action.

@@

\section{Grounding the purposiveness of joint action}
So far we have only been considering a possible role for social motor representation in facilitating joint action.  
Our primary concern, though, is with what joint action is.
The details of how 
social motor representation enables joint action
 already give us grounds for holding that motor representation has a role to play in explaining which events are joint actions.
To see why,
let’s go back to individual action for a moment again.

[***moved to earlier]

The same two questions we asked about ordinary, individual action also arise for joint action.
%What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?
What singles out the outcome or outcomes to which a purposive joint action is directed?
And what binds together the various activities (of several agents) that make up the joint action?
%A difference in the case of joint action is, of course, that the component activities are not activities of a single agent.

>>>>

\input{$HOME/Documents/submissions/preamble_steve_paper3}
%\author{}
%\date{}

@@

\section{***REFRAME***}
Shouldn't start with the question, Which events are joint actions?  Instead: either it's about making sense of purposive joint action, or else it's the question about the relation between purposive joint actions and the goals to which they are directed.
(Was just trying to make it sound more important by asking Which events are joint actions.  But I don't think this question bears any weight.)

\section{***EVIDENCE***}
I now think three bits of evidence are relevant:
\begin{enumerate}
\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).  This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
\end{enumerate}

@@

\subsection{My view}
Let us focus on the second consideration, 
which hinges on  a parallel between joint action and ordinary, individual action.
We take for granted that the parallel obtains but we do not accept the view about ordinary, individual action.

Not everyone accepts that for an event to be an action is for it to be appropriately related to an intention.
Some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' \citet[p.\ 251]{bratman:2000_valuing}.%
%\footnote{
%For instance, \citet[p.\ 251]{bratman:2000_valuing} % = \citep[p.\ 51]{Bratman:2007bd}
%describes an agent called `Creature 2' who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.
%}

@@

\section{Intention and Motor Representation}
\label{sec:individual_action_motor_representation}
[\emph{*Probably skip all of this section in the talk}]

@@

Motor representations are relevantly similar to intentions,
as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).%
\footnote{
Of course motor representations differ from intentions in some important ways (as these authors note).
But they are similar in the respects that matter for explaining the purposiveness of action.
}
A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.

How do motor representations resemble intentions?
First, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
So anyone who accepts the standard story about purposive action and intention
should also accept that a similar story about purposive action and motor representation.
Given that the two basic questions about ordinary, individual purposive action can also be answered by appeal to intention,
they can also be answering by appeal to motor representation.
This motivates the following view.
Not all purposive actions are bound together and linked to outcomes by intentions. 
In some cases what binds together purposive actions and links them to outcomes are motor representations.
%In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
And of course in many cases it may be that both intention and motor representation are involved.

This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining which events are actions (and plays a role resembling that of intention).
Now let's return to joint action.
%Suppose that the purposiveness of some actions really is grounded in motor representation in the way just described.
Could there be a parallel role for motor representation in explaining which events are joint actions?

@@

This needs unpacking. 
As already mentioned, a \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
To repeat, motor representations are not concerned with merely kinematic or dynamic features of actions only.
Rather, some motor representations represent outcomes, such as the movement of a target object from one place to another.

@@

Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome. 
It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
It is more controversial that reciprocal motor representation occurs in joint action.
Some of the research that Gunether Knoblich presented supports the claim that there are reciprocal motor representations in joint action, and he suggests that these enable joint action.%
\footnote{
See \citet{kourtis:2012_predictive}: `the partner’s expected action is simulated at the motor level, which probably facilitates effective performance of the joint action.'
\citet{kourtis:2010_favoritism} show that 
reciprocal motor representation is more likely to occur in joint action than is mere observation. 
See also \citet{Knoblich:2003nf}.
 }

I want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of each of the agents.
So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions.
For example, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
In this case I may represent the movement of the object from A to B and not only the component movements.
I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
%
%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.

@@

What follows is speculative philosophy: we take for granted that  social motor representation sometimes facilitates joint action and ask whether this conjecture bears on our question about which events are joint actions.

@@

To answer this question
let us take a step back and consider an individual action.
Suppose an agent moves a mug from one place to another, passing it from her left hand to her right hand half way.
It is a familiar idea that motor planning, like planning generally, involves starting with relatively abstract representations of outcomes and gradually filling in details.
We can capture this by supposing that 
motor representations for planning and monitoring action involve a hierarchical structure of representations.
At the top we might find a relatively abstract representation of an outcome, in this case of the movement of the object from one location to another.
Action-relevant details are progressively filled in by representations at lower stages of the hierarchy. 
Now in the action we are considering there is a need, even for the single agent, to coordinate the exchange between her two hands.
How is this achieved? 
We suppose that part of the answer involves the fact that planning for the movements of each hand is not done entirely independently.
Rather there is a plan for the whole action
and plans for the movements of each hand are components of this larger plan.
It is in part because they are parts of a larger plan that the plan for one hand constrains and is constrained by the plan for the other hand.

How is this relevant to the case of joint action?
In joint action the agents have the same goal, to move the object from one place to another.
They also face a similar coordination problem, requiring a precisely timed swap from one hand to another.
Now suppose, 
inspired by \citet{kourtis:2010_favoritism,kourtis:2012_predictive} and other findings mentioned above,
that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
The planning is the same almost up to the actual muscle contractions.

How could this be helpful?
Suppose the agents' planning processes are similar enough that, for a given context and problem, they will produce approximately the same plans.
Then having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action,
and 
(iii) each  agent's plans for their own action are constrained by the plans for the other agent's action.

@@

\section{Grounding the purposiveness of joint action}
So far we have only been considering a possible role for social motor representation in facilitating joint action.  
How does any of this bear on our main question about which events are joint actions?
The details of how 
social motor representation enables joint action
 already give us grounds for holding that motor representation has a role to play in explaining which events are joint actions.
To see why,
let’s go back to individual action for a moment again.

Earlier, in section \vref{sec:individual_action_motor_representation}, we asked two questions we asked about ordinary, individual action; the same two questions also arise for joint action.
%What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?
What singles out the outcome or outcomes to which a purposive joint action is directed?
And what binds together the various activities (of several agents) that make up the joint action?
%A difference in the case of joint action is, of course, that the component activities are not activities of a single agent.

### 2012-10-09 joint action and knowing others minds/interacting mindreaders.3.ps.corrections.tex
tidying up

Our question is part of a broader question,
What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
%\footnote{
%See e.g.\ \citet[pp.\ 126-7]{Davidson:1973jx}.
%}
%To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
%% nor directly about the mechanisms, processes or representations involved in mindreading
%, Davidson 
%sometimes formulates the question by asking what someone \emph{could} know that \emph{would} put them in a position to identify another's thoughts and actions.%
%\footnote{
%See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
%While what follows draws  on Davidson's questions and insights,
%our aims here are more modest than his.
%For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
%and we are not concerned with larger claims about mind, meaning and truth 
%(on which see 
%\citet{Davidson:1990du} and 
%\citet{lepore_donald_2005}).
%}
The most sustained attempts to answer this question,
Davidson's (\citeyear{Davidson:1984wh}; \citeyear{Davidson:1990du}), Lewis' (\citeyear{lewis:1974ri}) and Dennett's (\citeyear{Dennett:1987sf}),
do not exploit the possibility of interaction.
The evidence and principles they consider are available to manifestly passive observers.
So on their theories,
a purely passive mindreader observing from behind a one-way mirror
is on a par with
a mindreader who, actually or apparently, could interact with those she seeks to interpret.
The two are on a par in this sense:
in principle the same evidence could be available to each, and each can exploit the routes to knowledge in moving from evidence to ascriptions of thought and action.
Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
But on these theories interaction makes no difference to what can in principle be known.
In this paper we aim to show
that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
Their ascriptions could exploit routes to knowledge which would be unavailable if they were entirely passive observers.

@@

\end{enumerate}
%
Therefore:
%
\begin{enumerate}[resume]
%
\item A goal of your actions will be my goal, the goal I now envisage that my actions will be directed to.
\end{enumerate}
%
Call this inference \emph{your-goal-is-my-goal}.  
To say that it characterises a route to knowledge implies two things.  
First, in some cases it is possible to know the premises, 1–2, without already knowing the conclusion, 3.  
Second, in some of those cases knowing the premises would put one in a position to know the conclusion.  
We shall consider these points in turn

>>>>

Our question is part of a broader question,
What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
%\footnote{
%See e.g.\ \citet[pp.\ 126-7]{Davidson:1973jx}.
%}
%To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
%% nor directly about the mechanisms, processes or representations involved in mindreading
%, Davidson 
%sometimes formulates the question by asking what someone \emph{could} know that \emph{would} put them in a position to identify another's thoughts and actions.%
%\footnote{
%See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
%While what follows draws  on Davidson's questions and insights,
%our aims here are more modest than his.
%For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
%and we are not concerned with larger claims about mind, meaning and truth 
%(on which see 
%\citet{Davidson:1990du} and 
%\citet{lepore_donald_2005}).
%}
The most sustained attempts to answer this question,
Davidson's (\citeyear{Davidson:1984wh}; \citeyear{Davidson:1990du}), Lewis' (\citeyear{lewis:1974ri}) and Dennett's (\citeyear{Dennett:1987sf}),
do not exploit the possibility of interaction.
The evidence and principles they consider are available to manifestly passive observers.
So on their theories,
a purely passive mindreader observing from behind a one-way mirror
is on a par with
a mindreader who, actually or apparently, could interact with those she seeks to interpret.
The two are on a par in this sense:
in principle the same evidence could be available to each, and each can exploit the 
%correction! ***
same 
routes to knowledge in moving from evidence to ascriptions of thought and action.
Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
But on these theories interaction makes no difference to what can in principle be known.
In this paper we aim to show
that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
Their ascriptions could exploit routes to knowledge which would be unavailable if they were entirely passive observers.

@@

\end{enumerate}
%
Therefore:
%
\begin{enumerate}[resume]
%
\item A goal of your actions will be my goal, the goal I now envisage that my actions will be directed to.
\end{enumerate}
%
Call this inference \emph{your-goal-is-my-goal}.  
To say that it characterises a route to knowledge implies two things.  
First, in some cases it is possible to know the premises, 1–2, without already knowing the conclusion, 3.  
Second, in some of those cases knowing the premises would put one in a position to know the conclusion.  
We shall consider these points in turn. %correction! (the period) ***

### 2012-10-09 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
tidying up

But there is also a body of evidence supporting a different conclusion.
\citet{back:2010_apperly} found that subjects are significantly slower to answer an unexpected question about another's true or false belief compared with matched questions about reality \citep[see also][]{apperly:2006_belief}.
This suggests that, at least in adults, belief tracking is not automatic.
There is also evidence that, even in relatively simple situations, 
using facts about others' beliefs is not automatic \citep{Keysar:2003xu,apperly:2010_limits}.
The case for nonautomaticity is indirectly supported by evidence that tracking perceptions and beliefs---and even merely holding in mind what another believes, where no inference is required---involves a measurable processing cost  \citep{apperly:2008_back,apperly:2010_limits}, consumes attention and working memory in fully competent adults (\citealp{en_1698, lin:2010_reflexively, en_1547} experiments 4-5), may require inhibition \citep{bull:2008_role} and makes demands on executive function \citep{apperly:2004_frontal,samson:2005_seeing}.
These findings, taken together, suggest that tracking others' perceptions and beliefs is sometimes not automatic.

@@

\subsection{Extensions and variations}
With the fourth principle we have completed the construction of a minimal theory of mind capable of underwriting success on some false belief tasks.  
We stop here because false belief tasks are often taken to be an acid test for theory of mind.
But of course additonal principles can be added accommodate further theory of mind abilities.  
For instance, further principles might extend the definition of registration.  
Registration was been defined as a relation between agents, objects and locations. 
This definition could be extended to include  other types of property as relata in addition to locations.  
Further notions, such as a relational proxy for tracking desires, could also be added.  
These and other modifications would enable hypotheses about minimal theory of mind cognition  to explain a wider range of theory of mind abilities.%
\footnote{
Infants and two-year-olds' theory of mind abilities involve more than tracking false beliefs about location.
To illustrate, \citet{en_1691} show that infants can use communicated information in predicting actions based on false belief, 
\citet{he:2011_false}  show that 2.5-year-olds can succeed on false belief tasks involving unexpected contents rather than location properties, 
and  \citet{en_1795}  show that 18-month olds are sensitive to inferences that others are likely to make. 
}

>>>>

But there is also a body of evidence supporting a different conclusion.
\citet{back:2010_apperly} found that subjects are significantly slower to answer an unexpected question about another's true or false belief compared with matched questions about reality \citep[see also][]{apperly:2006_belief}.
This suggests that, at least in adults, belief tracking is not automatic.
There is also evidence that, even in relatively simple situations, 
using facts about others' beliefs is not automatic \citep{Keysar:2003xu,apperly:2010_limits}.
The case for nonautomaticity is indirectly supported by evidence that tracking perceptions and beliefs---and even merely holding in mind what another believes, where no inference is required---involves a measurable processing cost  \citep{apperly:2008_back,apperly:2010_limits}, consumes attention and working memory in fully competent adults (\citealp{Apperly:2009cc, lin:2010_reflexively, McKinnon:2007rr} experiments 4-5), may require inhibition \citep{bull:2008_role} and makes demands on executive function \citep{apperly:2004_frontal,samson:2005_seeing}.
These findings, taken together, suggest that tracking others' perceptions and beliefs is sometimes not automatic.

@@

\subsection{Extensions and variations}
With the fourth principle we have completed the construction of a minimal theory of mind capable of underwriting success on some false belief tasks.  
We stop here because false belief tasks are often taken to be an acid test for theory of mind.
But of course additional principles can be added accommodate further theory of mind abilities.  
For instance, further principles might extend the definition of registration.  
Registration was been defined as a relation between agents, objects and locations. 
This definition could be extended to include  other types of property as relata in addition to locations.  
Further notions, such as a relational proxy for tracking desires, could also be added.  
These and other modifications would enable hypotheses about minimal theory of mind cognition  to explain a wider range of theory of mind abilities.%
\footnote{
Infants and two-year-olds' theory of mind abilities involve more than tracking false beliefs about location.
To illustrate, \citet{en_1691} show that infants can use communicated information in predicting actions based on false belief, 
\citet{he:2011_false}  show that 2.5-year-olds can succeed on false belief tasks involving unexpected contents rather than location properties, 
and  \citet{en_1795}  show that 18-month olds are sensitive to inferences that others are likely to make. 
}
